Effectful Semantics for Substructural SSA

### 2023-11-02 pldi24/paper.tex
Defined substructural effectful category

Note that we cannot simply have \(\ms{Rel}^{\{\ms{a}\}} = \ms{Rel}\), since, for any morphism \(R: \ms{Rel}(A, B)\) such that \(\exists a \in A, R(a) = \varnothing\), we have
\begin{equation}
  (R;\ms{drop}(B))(a) = \varnothing \neq \ms{drop}(A)(a) = \{()\}
\end{equation}
We will see that this corresponds to the fact that the following programs are not semantically equal
\begin{equation}
  \ms{let}\;x = \ms{loop}(); e \not\simeq e \qquad \text{where} \qquad x \notin \ms{fv}(e)
\end{equation}
Similarly, we cannot simply have \(\ms{Rel}^{\{\ms{r}\}} = \ms{Rel}\), since, for any morphism \(R: \ms{Rel}(A, B)\) such that \(\exists a \in A, |R(a)| \supseteq \{b_0, b_1\}\) where \(b_0 \neq b_1\), we have that
\begin{equation}
  (R;\ms{split}(B))(a) = \{(b, b) \mid b \in R(a)\} \neq (\ms{split}(A);R)(a) = R(a) \times R(a) 
\end{equation}
since only the left-hand side contains the pair \((b_0, b_1)\), this corresponding to the fact that the following programs are not semantically equal
\begin{equation}
  \ms{let}\;x = \ms{rand}(); (x, x) \not\simeq (\ms{rand}(), \ms{rand}())
\end{equation}
Rather than directly give a concrete example of a general substructural monoidal category, it makes more sense to first jump to the next definition, which will provide us with one more naturally:
\begin{definition}[Susbtructural Effectful Category]
  \ \TODO{this}
\end{definition}

>>>>

Note that we cannot simply have \(\ms{Rel}^{\{\ms{a}\}} = \ms{Rel}\), since, for any morphism \(R: \ms{Rel}(A, B)\) such that \(\exists a \in A, R(a) = \varnothing\), we have
\begin{equation}
  (R;\ms{drop}(B))(a) = \varnothing \neq \ms{drop}(A)(a) = \{()\}
\end{equation}
We will see that this corresponds to the fact that the following programs are not semantically equal
\begin{equation}
  \ms{let}\;x = \ms{loop}(); e \not\simeq e \qquad \text{where} \qquad x \notin \ms{fv}(e)
\end{equation}
Similarly, we cannot simply have \(\ms{Rel}^{\{\ms{r}\}} = \ms{Rel}\), since, for any morphism \(R: \ms{Rel}(A, B)\) such that \(\exists a \in A, |R(a)| \supseteq \{b_0, b_1\}\) where \(b_0 \neq b_1\), we have that
\begin{equation}
  (R;\ms{split}(B))(a) = \{(b, b) \mid b \in R(a)\} \neq (\ms{split}(A);R)(a) = R(a) \times R(a) 
\end{equation}
since only the left-hand side contains the pair \((b_0, b_1)\), this corresponding to the fact that the following programs are not semantically equal
\begin{equation}
  \ms{let}\;x = \ms{rand}(); (x, x) \not\simeq (\ms{rand}(), \ms{rand}())
\end{equation}
Rather than directly give a concrete example of a general substructural monoidal category, it makes more sense to first jump to the next definition, which will provide us with one more naturally:
\begin{definition}[Susbtructural Effectful Category]
  A \textbf{substructural effectful category} \(\mc{C}_1 \to \mc{C}_0\) consists of
  \begin{itemize}
    \item A substructural monoidal category \(\mc{C}_1\)
    \item A symmetric premonoidal category \(\mc{C}_0\)
    \item An identity-on-objects functor \(\upg{\cdot}{}: \mc{C}_1^\varnothing \to \mc{C}_0\) preserving all symmetric premonoidal structure (which naturally induces functors \(\upg{\cdot}{}: \mc{C}_1^q \to \mc{C}_0\) which do the same)
  \end{itemize}
  We'll define \(\mc{C}_0^q\) to simply be \(\mc{C}_0\) to allow us to write \(\mc{C}_p^q\) for all \(p \in \{0, 1\}\), \(q \subseteq \{\ms{a}, \ms{r}\}\). Following this convention, we define \(\upg{f}{0}\) to be \(\upg{f}{}\) for morphisms in \(\mc{C}_1^q\) and simply \(f\) for morphisms in \(\mc{C}_0\), and \(\upg{f}{1}\) to simply be the identity functor on \(\mc{C}_1^q\).
\end{definition}
